sinterest_rates <- import('https://raw.githubusercontent.com/Camila-RV/VieiraKhanna_Assignment3/master/data_raw/interest_rates_eu.csv')---
title: "MPP-E1180: Collaborative Social Data Analysis Assignment 3"
author: "Camila Vieira & Tarun Khanna"
date: "11 November 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pkgs <- c('dplyr', 'ggplot2')
```

## Collected Indicators and Hypotheses

Source 1: WDI
- GDP per capita (constant 2000 US$)
Definition: Gross domestic product divided by midyear population.
Hypothesis: Economic growth and income have an important and positive impact on green investment.  They generate a higher demand for energy and clean air as well as higher economic means for adopting new technologies.

((((- Profit tax (% of commercial profits)
Definition: Profit tax is the amount of taxes on profits paid by the business.
Hypothesis: Increases of taxes on business may negatively impact companies capital margin for investing in installation capacity of renewable energy sources.)))) 
  
- Energy imports, net (% of energy use)
Definition: Net energy imports are estimated as energy use less production, both measured in oil equivalents. Energy use refers to use of primary energy before transformation to other end-use fuels, which is equal to indigenous production plus imports and stock changes, minus exports and fuels supplied to ships and aircraft engaged in international transport.

Hypothesis: The energy imports indicator may allows us to verify whether countries have reduced dependence on energy sources from other countries by investing in renewable installations.
We expect that a lower energy imports value increases the likelihood of countries investing in the installation of renewable energy sources. 
  
  /// If a country presents a negative value it indicates that the country is a net exporter. ->> How to address this point? ///

(((- Fossil fuel energy consumption (% of total)
Definition: Use of energy from fossil fuel as a percent of total energy use. Fossil fuel comprises coal, oil, petroleum, and natural gas products.
Hypothesis: The greater the percentage of energy consumption is the lower will be the incentives for developing installation capacity of renewable energy sources through new investments. ))))

Source 2: Eurostat

- Innovation in Renewable Energy
Definition: Number of patents filed for renewable energy/climate mitigation country wise. It comprises energy technologies patent applications submitted to the European Patent Office.
Hypothesis: The number of patent applications in renewable energy is used as a proxy for innovation levels having a positive impact on investments in renewable resources. 

Source 3: European Central Bank

- Long-term interest rate (10 ys. maturity)
Definition:

Hypothesis:

- Oil Price


```{r cars}

source('Eurostat_data.R')
source('WDI_data.R')
data <- merge(Combined_WDI,Combine_EuroStat,by = c("country", "year"), all.x = T)



```



```{r pressure, echo=FALSE}



```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
