---
title: 'Determinants of Renewable Energy: An EU Cross-Country Analysis'
subtitle: 'MPP-E1180: Collaborative Social Science Data Analysis'
author: Camila Vieira & Tarun Khanna
date: December 2nd, 2016
output: 
  ioslides_presentation
transition: faster
smaller: true 

---

## Solar and Wind energy penetration in Europe

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, type = 'html', results='asis'}

setwd("~/Desktop/CollaborativeDataMethods/VieiraKhanna_Assignment3/data_cleaning")

# Importing and cleaning data from Eurostat
source('Eurostat_and_Opec_data.R')

# Importing and cleaning data from World Bank
source('WDI_data.R')

# Merging the data
re_data <- merge(Combined_WDI,Combine_EuroStat,by = c("country", "year"), all.x = T, all.y = T)

# Remove missing values
re_data1 <- na.omit(re_data)

#creating var for solar and wind gen as percent of total elec consumption
re_data1$sw_pc_elec <- re_data1$re_gen/re_data1$elec_total*100 
re_data_map <- re_data1[which(re_data1$year == "2013"),]

library(googleVis)
# Plot
re_map <- gvisGeoChart(re_data_map, locationvar = 'country',
                      colorvar = 'sw_pc_elec',
                      options = list(
                        colors = "['#f7fcb9', '#31a354']",
                        region ="150"
                          ))

print(re_map, tag = 'chart')

```

The map shows the proportion of total electricity consumption attributable to wind and solar in Europe in 2013

## Trends in Wind Electricity Generation 

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, fig.width = 6, fig.height = 4}

# Data visualisation
library(ggplot2)
library(dplyr)

#selecting data for wind
re_data1 <- re_data1[order(re_data1$year,re_data1$wind_gen),] #arranging data
countries10 <- tail(re_data1,10)[1] %>% as.matrix() %>% as.character() %>% paste(collapse = "|") #finding top10
re_wind_top10  <- filter(re_data, grepl(countries10, country)) #selecting top 10 countries by generation 

#create graph of Wind Generation
Colours <- c('#543005','#8c510a','#bf812d','#dfc27d','#f6e8c3','#c7eae5','#80cdc1',
             '#35978f','#01665e','#003c30') # assigning colour pallette

ggplot(data = na.omit(re_wind_top10)) + 
  aes(x = year, y = wind_gen, group = country, fill = country) +
  geom_area(position = "stack") + 
  scale_color_manual(values = Colours) +
  xlab("\nYear") +
  ylab("Wind Energy Generation(TWh)\n") +
  ggtitle("Wind Energy Generation in ten largest wind energy markets in Europe\n") +
  theme_bw(base_size = 10)

rm(list = c("re_wind_top10", "countries10"))
```

- Wind energy generation in top 10 markets quadrupled 
- Germany, Spain and UK are the largest markets by size for wind

## Trends in Solar Electricity Generation

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, fig.width = 6, fig.height = 4}

#selecting data for solar
re_data1 <- re_data1[order(re_data1$year,re_data1$solar_gen),] #arranging data
countries10 <- tail(re_data1,10)[1] %>% as.matrix() %>% as.character() %>% paste(collapse = "|") #finding top10
re_solar_top10  <- filter(re_data, grepl(countries10, country)) #selecting top 10 countries by generation 

#create graph of Solar Generation
ggplot(data = na.omit(re_solar_top10)) + 
  aes(x = year, y = solar_gen, group = country, fill = country) +
  geom_area(position = "stack") + 
  scale_color_manual(values = Colours) +
  xlab("\nYear") +
  ylab("Solar Energy Generation (TWh)\n") +
  ggtitle("Solar Energy Generation in ten largest solar energy markets in Europe\n") +
  theme_bw(base_size = 10)

rm(list = c("re_solar_top10", "countries10","p","countries","Colours"))
```

- Meteoric growth in solar energy generation
- Largest markets for solar energy are Germany, Spain and Italy

## Our Research Question  

__What are the determinants of renewable energy in EU countries?__  
<span style="color:white">.</span>  
We try to answer this question using data for 28 EU-countries (2005-2013). Our data was sourced from Eurostat, The World Bank and OPEC and is also available on our GitHub repository.

## Our Analysis

### Dependent Variable:

Percentage of total energy consumption coming from solar and wind (sw pc elec)

### Independent variables:
- GDP per Capita ('000 USD)
- Net Energy Imports as a percentage of total energy consumption (%)
- Number of patents filed in RE technologies   
- Long-term interest rates (%, 10 Year Bond Yield)
- Crude oil prices (USD)

## Model
- We estimate the model: $sw pc elec = \alpha + \beta1*GDP capita + \beta2*net energy imports$   
$+ \beta3*interest rates + \beta4*oil_price + \beta5*patents$
- Different model specifications: pooled ordinary least squares, fixed effects model, random effects model
- The Hausman test indicates presence of panel specific effects 

## Regression results

```{r, include=F, message=F, warning=FALSE, error=FALSE, echo=FALSE}

library(plm) #for panel data
library(ggplot2) #for graphs using ggplot

# estimate pooled OLS model
m.pols  <- plm(sw_pc_elec  ~ log(GDP_capita) + netenergy_imports + interest_rate +  oil_price + poly(patents,3), data= re_data1, index=c("country", "year"), model="pooling")
summary(m.pols)

#re_data2 <- re_data1[-which(re_data1$country == "Luxembourg"),]

# estimate panel fixed effects 
m.fe  <- plm(sw_pc_elec ~ log(GDP_capita) + netenergy_imports  + interest_rate +  oil_price + patents, data= re_data1, index=c("country", "year"), model="within")
summary(m.fe)

#re_data2 <- re_data1[-which(re_data1$country == "Luxembourg"),]
#m.fe1  <- plm(re_gen  ~ log(GDP_capita) + netenergy_imports  + interest_rate +  oil_price + poly(patents,3), data= re_data2, index=c("country", "year"), model="within")
#summary(m.fe1)

```

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, fig.width = 6, fig.height = 4}

library(sjPlot)
library(sjmisc)

# graphs for presentation
sjp.lm(m.fe, title = "", theme_bw()) # plotting coeffificents on graph

```

_** (1) logged GDP has a statistically significant negative coefficient (2) the model has adjusted R2 = 43%_

## Effect of Oil Prices

```{r, include=F, message=F, warning=FALSE, error=FALSE, echo=FALSE}

#Modeling interactions
m.fe.i  <- plm(sw_pc_elec  ~ oil_price*netenergy_imports + log(GDP_capita) + patents + interest_rate, data= re_data1, index=c("country", "year"), model="within")
summary(m.fe.i)

vcov(m.fe.i) #variance covariance matrix

vb1 <- vcov(m.fe.i)[1]
vb3 <- vcov(m.fe.i)[6,6]
cvb1b3 <- vcov(m.fe.i)[1,6]

margins <- data.frame(
  netenergy_imports = min(re_data1$netenergy_imports, na.rm =T):max(re_data1$netenergy_imports, na.rm =T),
  me = coef(m.fe.i)[1] + coef(m.fe.i)[6] * ( 
  min(re_data1$netenergy_imports, na.rm =T):max(re_data1$netenergy_imports, na.rm =T)),
  se = sqrt(vb1 + (min(re_data1$netenergy_imports, na.rm =T):max(re_data1$netenergy_imports, na.rm =T))^2 * vb3 + 2 *(min(re_data1$netenergy_imports, na.rm =T):max(re_data1$netenergy_imports, na.rm =T)) * cvb1b3)
)
margins$lwr <- margins$me - 1.96 * margins$se
margins$upr <- margins$me + 1.96 * margins$se
```

#### Marginal Effect of Oil Price Conditioned Over Import Intensity of Energy

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, fig.width = 6, fig.height = 4}
ggplot(data = margins, aes(x = netenergy_imports, y = me, ymin = lwr, ymax = upr)) +
  geom_ribbon(fill = 'blue', alpha = 0.2) + 
  geom_line(colour = "red", size =1) + theme_bw() + 
  geom_hline(yintercept = 0, linetype = 'dashed') + 
  xlab("\nEnergy Imports (% of consumption)") + 
  ylab("Marginal Effect of Oil Price\n") +
  ggtitle("") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'))

```

## Country Specific Effects 

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, eval= T}

country_effect <- data.frame(summary(fixef(m.fe))[,1],summary(fixef(m.fe))[,4]) 
country_effect$country <- rownames(country_effect)
names(country_effect)[1] <- "intercept"
names(country_effect)[2] <- "pvalue"
country_effect$prop_intercept <- country_effect$intercept/min(country_effect$intercept)

#calculating FE of each country relative to min
country_effect <- arrange(country_effect,prop_intercept)
countries10 <- tail(country_effect,10)[3] %>% as.matrix() %>% as.character() %>% paste(collapse = "|") #finding top10
country_effect10  <- filter(country_effect, grepl(countries10, country)) #selecting top 10 countries by fixed effect 

# graphing relative FE
ggplot(data = country_effect10) + 
  aes(country, prop_intercept) + 
  geom_bar(stat = "identity") + 
  coord_cartesian(ylim = c(1,1.75)) +  
  xlab("\n Country") +
  ylab(" ") +
  ggtitle("Proportionate Country Effect relative to minimum\n") +
  theme_bw(base_size = 10) 
```

Country specific effects are statistically significant

## Overall Results and Conclusion

- Renewable energy uptake varies significantly across EU countries 
- The increase in cost of conventional energy seems to be an important factor in explaining 
RE development
- Country specific conditions including renewable energy development policy, geography plays an important role
- Next steps: disentangle the effect of policy from country specific effects  
